%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% LIVECOMS ARTICLE TEMPLATE FOR BEST PRACTICES GUIDE
%%% ADAPTED FROM ELIFE ARTICLE TEMPLATE (8/10/2017)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PREAMBLE
\documentclass[9pt,tutorial]{livecoms}
% Use the 'onehalfspacing' option for 1.5 line spacing
% Use the 'doublespacing' option for 2.0 line spacing
% Use the 'lineno' option for adding line numbers.
% Use the "ASAPversion' option following article acceptance to add the DOI and relevant dates to the document footer.
% Use the 'pubversion' option for adding the citation and publication information to the document footer, when the LiveCoMS issue is finalized.
% The 'bestpractices' option for indicates that this is a best practices guide.
% Omit the bestpractices option to remove the marking as a LiveCoMS paper.
% Please note that these options may affect formatting.

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx,bm}
\usepackage{hyperref} % Not original - Added by Jack
\usepackage{listings} % Not original - Added by Jack
\DeclareSIUnit\Molar{M}
\usepackage[italic]{mathastext}
\newcommand{\pka}{p\textit{K}$_{\rm a}$}
\graphicspath{{figures/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% IMPORTANT USER CONFIGURATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\versionnumber}{1.0}  
% you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/JanaShenLab/CpHMD-Tutorial}}  %this should be the main github repository for this article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{A Guide to the Continuous Constant pH Molecular Dynamics 
Methods in Amber and CHARMM Article [v\versionnumber]}

\author[1,\authfn{1},\authfn{2}]{Jack A. Henderson}
\author[1,\authfn{1}]{Ruibin Liu}
\author[1,\authfn{3}]{Julie A. Harris}
\author[1,\authfn{4}]{Yandong Huang}
\author[1]{Vinicius Martins de Oliveira}
\author[1*]{Jana Shen}
\affil[1]{University of Maryland School of Pharmacy, Baltimore, MD}
%\affil[2]{Institution 2}

\corr{jana.shen@rx.umaryland.edu}{JS} 

\orcid{Jack A. Henderson}{https://orcid.org/0000-0001-6675-7944}
\orcid{Ruibin Liu}{https://orcid.org/0000-0001-8395-9353}
\orcid{Julie A. Harris}{https://orcid.org/0000-0002-1130-6457}
\orcid{Yandong Huang}{https://orcid.org/0000-0002-1452-6383}
\orcid{Vinicius M. de Oliveira}{https://orcid.org/0000-0003-0927-3825}
\orcid{Jana Shen}{https://orcid.org/0000-0002-3234-0769}

\contrib[\authfn{1}]{These authors contributed equally to this work}

\presentadd[\authfn{2}]{Scorpion Therapeutics, Boston, MA}
\presentadd[\authfn{3}]{ComputChem LLC, Baltimore, MD}
\presentadd[\authfn{4}]{Jimei University, College of Computer Engineering, Xiamen, China}

\blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository. To provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PUBLICATION INFORMATION
%%% Fill out these parameters when available
%%% These are used when the "pubversion" option is invoked
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pubDOI{10.33011/livecoms.4.1.1563}
\pubvolume{4}
\pubissue{1}
\pubyear{2022}
\articlenum{1563}
\datereceived{22 January, 2022}
\dateaccepted{6 August, 2022}

% \lstnewenvironment{listings}[1][] 
%  {\lstset{frame=off,escapechar=`,breaklines, breakindent=10pt, breakatwhitespace=True, linewidth=3.4in, #1}}
% {}

\lstnewenvironment{plaincontent}[1][] 
 {\lstset{frame=shadowbox,escapechar=`,breaklines, breakindent=10pt, breakatwhitespace=True, linewidth=3.4in, #1}}
{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}
\maketitle

\begin{abstract}
Like temperature and pressure, solution pH is an important environmental variable in biomolecular simulations. 
Virtually all proteins depend on pH to maintain their structure and function. 
In conventional molecular dynamics (MD) simulations of proteins, pH is implicitly accounted for by assigning and fixing protonation states of titratable sidechains. This is a significant limitation, 
as the assigned protonation states may be wrong and they may change during dynamics.
In this tutorial, we guide the reader in learning and using the various continuous constant pH MD methods in Amber and CHARMM packages, which have been applied to predict {\pka} values and elucidate proton-coupled conformational dynamics of a variety of proteins including enzymes and membrane transporters.
\end{abstract}

\end{frontmatter}

\input{1_intro}
\input{1_methods}
\input{1_scope}
\input{2_prereq}
\input{3_hphmd}
\input{4_aphmd}
\input{5_gphmd}
\input{6_analysis}
\input{7_parameterization}

\section{Author Contributions}
Henderson, Liu, Harris, Huang, de Oliveira, and Shen
together wrote the manuscript.
We thank Zhi Yue and Paween Mahinthichaichan for contributing example scripts for the hybrid-solvent CpHMD simulations of transmembrane proteins.
We thank Jason A. Wallace for a bash script to calculate {\pka} values and track the convergence of unprotonated fractions using Grace.

\section{Potentially Conflicting Interests}
%%%%%%%
%Declare any potentially competing interests, financial or otherwise
%%%%%%%
The CpHMD-related software may be of interest to ComputChem LLC, for which
J.S. is a founder and scientific advisor.
J.S. also serves on the scientific advisory board of MatchPoint Therapeutics.

\section{Funding Information}
%%%%%%%
% Authors should acknowledge funding sources here. Reference specific grants.
%%%%%%%
We acknowledge financial support from the National Institutes of Health (R01GM098818 and R01CA256557 to Shen and R44GM134756 to Harris) and National Science Foundation (CBET1932963 to Shen). 
We thank Zhi (Shane) Yue for contributing the input files for the membrane-enabled hybrid-solvent CpHMD simulations.
We thank Paween Mahinthichaichan for providing comments on the input files for the membrane-enabled hybrid-solvent CpHMD simulations.


\section*{Author Information}
\makeorcid

\bibliography{cphmd}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\appendix


\end{document}
